---
title: "RF profiling in Sjögren’s disease with lymphoma"
subtitle: "Rheumatology · Manuscript RHE-25-1792.R2"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
    theme: flatly
    self_contained: true
---

This notebook is the single source of executable analysis. It starts from the
curated inputs used by the supplied script and regenerates figures and tables.
It does not rerun raw proteomics processing, structural docking or mCSM-ab2.
After running the notebook, open the generated HTML to read the results; use the
**Code** buttons to inspect each calculation. To rerun, follow `README.md`.

## Environment and inputs

```{r setup}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE,
  error = FALSE, fig.width = 10, fig.height = 6, dpi = 150, dev.args = list(bg = "white"))
options(scipen = 999, digits = 4)
set.seed(1792)
dir.create("outputs", showWarnings = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(forcats)
library(purrr)
library(tibble)
library(ggplot2)
library(rstatix)
library(ggpubr)
library(MetBrewer)
library(cowplot)
library(egg)
library(ggh4x)
library(ggmsa)
library(seqinr)
library(ggbeeswarm)
library(scales)
# Resolve masked functions explicitly, independently of the user's workspace.
select <- dplyr::select
filter <- dplyr::filter
rename <- dplyr::rename
count <- dplyr::count
desc <- dplyr::desc
alpha <- ggplot2::alpha
ggarrange <- ggpubr::ggarrange

```

```{r check-inputs}
alignment_lines <- read_lines("data/rf_ighv_peptide_alignment.txt")
stopifnot(length(alignment_lines) <= 999,
          trimws(alignment_lines[350]) == "Non-lymphoma")
```

## 1. Decode peptide alignments

Spaces are unobserved; dots match the reference; letters are residues. Explicit reference-matching letters are dropped by the parser.

```{r analysis-1}
get_contig_no<-function(string_of_aa_positions) {
  length(string_of_aa_positions)->l
  x = 2
  s = 1
  section = c(s)
  while (x <= l) {
    if (string_of_aa_positions[x]-string_of_aa_positions[x-1] == 1) {
      section = append(section, s)
    } else {
      s = s + 1
      section = append(section, s)
    }
    x = x + 1
  }
  return(section)
}

raw_seq_parser<-function(data_column, spacer_symbol,imgt_ref_long) {
  data_raw<-na.omit(data_column)
  arranged_table<-tibble()
  for (i in 1:length(data_raw)) {
    x<-data_raw[i]
    
    pt_code<-if_else(
      str_detect(x,"^\\d"),
      str_extract(x, "\\d+-\\d+"),
      NA_character_
    )
    
    sf<-if_else(str_detect(x,">"),str_extract(x,"(?<=IGHV)[\\d-_]+(?=|)") %>% str_to_lower(),NA_character_)
    
    string<-if_else(!str_detect(x,">|\\d"),x,NA_character_)
    
    arranged_table<-arranged_table %>% 
      bind_rows(tibble(
        read_unique_id=i,
        pt_code=pt_code,
        subfamily = sf,
        string=string
      ))
  }
  arranged_table<-arranged_table %>%
    fill(c(pt_code, subfamily),.direction = "down") %>%
    filter(!is.na(string))%>%
    separate(
      string,
      sep = "",
      into = paste0(seq.int(0,108,1))) %>%
    select(-`0`) %>%
    pivot_longer(
      cols = c(`1`:`108`),
      names_to = "aa_position",
      values_to = "aa",
      values_transform = list(
        aa = as.character,
        aa_position = as.numeric
      )
    ) %>%
    filter(aa != spacer_symbol,!is.na(aa), aa != "") %>%
    mutate(
      subfamily = str_replace(subfamily,"_","*"),
      aa = str_to_upper(aa),
      aa_position = as.numeric(aa_position)) %>%
    left_join(imgt_ref_long) %>%
    filter(aa != aa_ref) %>% 
    group_by(read_unique_id) %>% 
    mutate(peptide_unique_id = paste0(read_unique_id,"-", get_contig_no(aa_position)))
  
  return(arranged_table)
}

pipe_print<-function(x) {
  print(x)
  return(x)
}
```

## 2. Reference numbering

Map the first 108 IMGT alignment columns to sequential residue positions.

```{r analysis-2}
read_imgt_ighv_fasta<-function(fasta_file){
  imgt<-read.fasta(file = fasta_file,as.string = T,seqtype = "AA") %>% 
    as_tibble()%>% t() %>% 
    as_tibble(rownames = NA)
  
  imgt_rownames<-rownames(imgt)
  
  imgt_ref<-bind_cols(imgt_rownames,imgt) %>% as_tibble(rownames = NULL)
  colnames(imgt_ref)<-c("ighv", "ref_seq")
  
  imgt_ref_long<-
    imgt_ref %>% 
    mutate(subfamily = str_extract(ighv,"(?<=IGHV)[\\d[[:punct:]]]+(?=|)"))%>%
    separate(ref_seq,sep = "",into = paste0("aa_",seq.int(0,108,1))) %>% 
    pivot_longer(
      cols = dplyr::starts_with("aa"),
      values_to = "aa_ref",
      names_to = "imgt_pos",
      values_transform = list(
        aa_ref = as.character,
        position =as.numeric
      )) %>% 
    mutate(imgt_pos = str_extract(imgt_pos, "\\d+") %>% as.numeric(),
           is.gap = case_when(
             str_detect(aa_ref, "\\.")~ "GAP",
             TRUE ~ NA_character_
           ))%>% 
    dplyr::filter(imgt_pos !=0, !is.na(aa_ref), is.na(is.gap)) %>% 
    group_by(subfamily) %>% 
    mutate(aa_position = row_number()) %>% 
    select(-is.gap,-ighv)
  
  return(imgt_ref_long)
  
}

imgt_reference<-read_imgt_ighv_fasta("data/reference/imgt_ighv_reference.fasta")
```

## 3. Amino-acid properties

Use the supplied amino-acid properties to score substitutions.

```{r analysis-3}
aa_characteristics_short<-read_csv("data/reference/amino_acid_properties.csv", show_col_types = FALSE) %>% 
  select(1:4) %>% 
  mutate(
    Aromatic = case_when(
      Code %in% c("F", "W", "Y") ~ "Aromatic",
      TRUE ~ "Non-aromatic")
    ) 

aa_characteristics<- aa_characteristics_short%>% 
  mutate(
    Change_to = toString(Code),
  ) %>% 
  rename(
    Change_from = Code,
    Hydropathy_from = Hydropathy,
    Charge_from = Charge,
    Aromatic_from = Aromatic
  ) %>% 
  separate(Change_to,sep = ", ",into = paste0("col_",seq.int(1,20,1))) %>% 
  pivot_longer(
    cols = starts_with("col_"),
    values_to = "Change_to", 
    names_to = NULL) %>% 
  left_join(aa_characteristics_short[,2:5] %>% 
              rename(Hydropathy_to = Hydropathy,
                     Charge_to = Charge,
                     Aromatic_to = Aromatic),
            by = c("Change_to" = "Code")
            ) 

aa_change_matrix<-aa_characteristics %>% 
  mutate(
    hy_change = case_when(
      Hydropathy_from == Hydropathy_to ~ 0,
      Hydropathy_to == "moderate" | Hydropathy_from == "moderate" ~ 0.5,
      TRUE ~ 1
    ),
    ch_change = case_when(
      Charge_from == Charge_to ~ 0,
      Charge_to == "N" | Charge_from == "N" ~ 0.5,
      TRUE ~ 1
    ),
    ar_change = case_when(
      Aromatic_from == Aromatic_to ~ 0,
      TRUE ~ 1
    ),
    change_vector = sqrt(hy_change^2+ch_change^2+ar_change^2),
  ) %>% select(Change_from,Change_to,change_vector,hy_change,ch_change,ar_change)
```

## 4. Colour key

The small TSV defines the seven residue groups. Colours come from the pinned ggmsa installation; the original use of its internal palette is retained.

```{r analysis-4}
zappo_colors = tibble(color = ggmsa:::scheme_AA$Zappo_AA,
                      aa = rownames(ggmsa:::scheme_AA))

aa_zappo <- read_tsv("data/reference/zappo_amino_acid_groups.tsv", col_names = c("X1", "X2"), show_col_types = FALSE) %>%
  rename(group = X1, aa_ = X2) %>%
  separate(col = aa_,
           sep = "",
           into = paste0("aa", seq.int(1, 10, 1))) %>%
  pivot_longer(
    starts_with("aa"),
    values_drop_na = T,
    names_to = NULL,
    values_to = "aa"
  ) %>%
  filter(aa != "") %>%
  mutate(r_num = row_number()) %>%
  group_by(group) %>%
  mutate(start = min(r_num) - 0.5, end  = max(r_num) + 0.5) %>%
  left_join(zappo_colors)

color_scale <- aa_zappo %>%
  distinct(group, color)
color_scale_zappo <- color_scale$color
names(color_scale_zappo) <- color_scale$group

color_scale_zappo %>% as.character()

scale_zappo<- scale_discrete_manual(aesthetics = "fill", values = color_scale_zappo,name = "")
```

## 5. Study data and filtering

Filter curated alignments at >1 observation and >10% mutation coverage; retain zero-count participants in summaries.

```{r analysis-5}
rf_data<-tibble(data = read_lines("data/rf_ighv_peptide_alignment.txt")) %>% 
  mutate(data = str_replace_all(data,"\\s","@"))
cryo_patients<-c("003-25", "003-27", "003-40","07-06", "07-12", "08-09")

prelymphoma_patients<- c("01-32", "07-12", "07-35", "08-09", "15-24")
ritux_treated<-c("01-06", "03-16", "07-06", "07-09", "07-33")

rf_levels <- read_csv("data/rf_levels.csv", col_types = cols(pt = col_character(), rf_level = col_double()))
vh_dominant <- read_csv("data/dominant_clones.csv", col_types = cols(.default = col_character()))

results_shm<-raw_seq_parser(rf_data$data[1:349], spacer_symbol = "@",imgt_ref_long = imgt_reference) %>% 
  mutate(Lymphoma = "Lymphoma") %>% 
  bind_rows(raw_seq_parser(rf_data$data[350:999], spacer_symbol = "@",imgt_ref_long = imgt_reference)%>% 
              mutate(Lymphoma = "Non Lymphoma")
  )%>% 
  rowwise() %>% 
  mutate(pt = paste0(pt_code),
         subfamily = str_extract(subfamily, "\\d-\\d{1,2}"),
         .keep = "unused"
         ) %>% 
  ungroup() %>% 
  group_by(subfamily,Lymphoma) %>% 
  mutate(pts_per_group_and_subfamily = n_distinct(pt)) %>% 
  group_by(pt, subfamily, imgt_pos) %>% 
  mutate(coverage_count = n()) %>% 
  ungroup() %>% 
  mutate(Cryoglobulin = case_when(
    pt %in% cryo_patients ~ "Cryo",
    TRUE ~ "No Cryo"
    ),
  `Pre-Lymphoma` = case_when(
    pt %in% prelymphoma_patients ~ "Pre Lymphoma",
    TRUE ~ Lymphoma
  ),
  `Current Lymphoma` = case_when(
    `Pre-Lymphoma` == "Lymphoma" ~ "Yes",
    TRUE ~ "No"
  ),
  `Pre-Lymphoma` = fct(`Pre-Lymphoma`,levels = c("Lymphoma","Pre Lymphoma","Non Lymphoma")),
  `B cell depletion` = case_when(
    pt %in% ritux_treated ~ "Ritux",
    TRUE ~ "No Ritux"
  ))

summarised_sf_table = results_shm %>% 
  select(pt,subfamily,Lymphoma) %>% 
  distinct() %>% 
  group_by(pt) %>% 
  summarise(subfamilies = toString(subfamily),sf_count = n())

pt_summary <- summarised_sf_table %>% left_join(
  results_shm %>%
    distinct(
      pt,
      Lymphoma,
      Cryoglobulin,
      `Pre-Lymphoma`,
      `Current Lymphoma`,
      `B cell depletion`
    )) %>% 
  left_join(rf_levels) %>% 
  left_join(vh_dominant) %>% 
  mutate(
    dominant = dominant %>% str_to_lower(),
    dominant_sf = case_when(
      str_detect(dominant, "^dominant") ~ "1-69",
      str_detect(dominant, "biclonal") ~ subfamilies,
      TRUE ~ str_extract(dominant,"\\d-\\d+")
    )
  )

shm_changes<-results_shm %>% 
  select(-contains("unique_id")) %>% 
  filter(
    aa != ".",
  ) %>% 
  group_by(pt,subfamily,imgt_pos,aa) %>%
  mutate(specific_aa_mut_coverage = n()) %>%
  ungroup() %>% 
  group_by(pt,subfamily,imgt_pos) %>% 
  mutate(any_aa_mut_coverage = n()) %>% 
  ungroup() %>%
  mutate(
    specific_mut_proportion=specific_aa_mut_coverage/coverage_count,
    any_mut_proportion = specific_aa_mut_coverage/coverage_count
  ) %>% 
  left_join(aa_characteristics %>% 
              left_join(aa_change_matrix), 
            by = join_by(aa_ref == Change_from, aa == Change_to)
  ) %>% 
  mutate(
    Hydrophobic_change = case_when(
      Hydropathy_to == "moderate" & Hydropathy_from =="hydrophilic" ~ "hydrophobic",
      Hydropathy_to == "moderate" & Hydropathy_from =="hydrophobic" ~ "hydrophilic",
      Hydropathy_from != Hydropathy_to ~ Hydropathy_to,
      TRUE ~ "Unchanged"
    ) %>% str_to_sentence(),
    Charge_change = case_when(
      Charge_to == "N" & Charge_from =="-" ~ "Positive",
      Charge_to == "N" & Charge_from =="+" ~ "Negative",
      Charge_from != Charge_to ~ Charge_to,
      TRUE ~ "Unchanged"
    )%>% str_to_sentence(),
    Charge_change = case_when(
      Charge_change == "-" ~ "Negative",
      Charge_change == "+" ~ "Positive",
      Charge_change == "N" ~ "Neutral",
      TRUE ~ Charge_change
    )
  ) %>% 
  distinct()

complete_data<-pt_summary %>% 
  separate(
    subfamilies,
    into = paste0("_SF",seq.int(1,10,1)),
    sep = ","
  ) %>% 
  pivot_longer(
    cols = starts_with("_SF"),
    values_transform = str_trim,
    values_drop_na = T,
    values_to = "subfamily",
    names_to = NULL
  ) %>% 
  left_join(
    shm_changes,
  ) 

filtered_data <- complete_data %>%
  ungroup() %>%
  filter(
    specific_aa_mut_coverage > 1,
    specific_mut_proportion > 0.1
  ) %>%
  mutate(kept = 1L)

p_rf_level<-pt_summary %>% ggplot()+
  aes(x = `Pre-Lymphoma`, y = rf_level)+
  geom_boxplot(outliers = FALSE)+
  geom_beeswarm(aes(shape = interaction(Cryoglobulin,`B cell depletion`,sep = "+\n",drop = T)),alpha = 1,cex = 3,size=3)+
  geom_pwc(p.adjust.method = "bonferroni",label = "p.adj.signif",hide.ns = F)+
  egg::theme_article(base_size = 11)+
  facet_wrap(vars("Rheumatoid factor level"))+
  scale_shape_manual(values = c(20,17,18,15),name = "")+
  guides(shape = guide_legend(direction = "horizontal",nrow = 1))+
  ylab("IU/L")+
  theme(legend.position = 'bottom',
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
        legend.direction = "vertical")
p_rf_level

# Report qualifying counts separately from unfiltered records.
supplemental_data <- complete_data %>%
  ungroup() %>%
  group_by(pt, subfamily, rf_level, Lymphoma, `Pre-Lymphoma`,
           `B cell depletion`, Cryoglobulin) %>%
  summarise(unfiltered_record_count = n(), .groups = "drop") %>%
  left_join(filtered_data %>% ungroup() %>%
              count(pt, subfamily, name = "qualifying_mutation_count"),
            by = c("pt", "subfamily")) %>%
  mutate(qualifying_mutation_count = replace_na(qualifying_mutation_count, 0L))
stopifnot(sum(supplemental_data$qualifying_mutation_count) == nrow(filtered_data))
write_csv(supplemental_data, "outputs/supplemental_data_table.csv")

```

## 6. Figure 1 — RF levels and IGHV repertoire

RF levels (IU/L), IGHV subfamilies and repertoire proportions.

```{r analysis-6}
repertoire_p1<-results_shm %>% 
  select(pt,subfamily,`Pre-Lymphoma`, `B cell depletion`) %>% 
  distinct() %>% 
  ggplot()+
  aes(x = pt, y = fct_infreq(subfamily), group = `Pre-Lymphoma` , 
      fill = `B cell depletion`)+
  geom_tile(alpha = 1, 
            lwd = 0.7,height = 0.8, width = 0.8)+
  scale_fill_met_d(name = "Hokusai2")+
  geom_tile(
    data = pt_summary %>% 
      select(pt,dominant_sf,`Pre-Lymphoma`) %>% 
      separate(dominant_sf,into = paste0("dsf_",seq_len(3)), sep = ", ") %>%
      pivot_longer(cols = starts_with("dsf_"),names_to = NULL,values_to = "dominant_sf",values_drop_na = T) %>% 
      mutate(color_lab = "Dominant"),
      aes(
        x = pt, y = dominant_sf, color = color_lab,fill = "transparent"
      ), 
    fill = "transparent",
    height = 0.8, 
    width = 0.8, 
    linewidth = 1,show.legend = T,inherit.aes = T)+
  scale_color_manual(values = c("black"))+
  facet_grid(cols = vars(fct_inorder(`Pre-Lymphoma`)),scales = "free_x",space = "free_x")+
  egg::theme_article(base_size = 11)+
  guides()+
  background_grid()+
  theme(legend.position = "bottom",
        legend.title = element_blank(),
    axis.text.x = element_text(angle = 90), 
    axis.title.x = element_blank(),
    strip.text = element_text(size = 10,face = "bold"),
    )+
  ylab("IGHV Subfamily")

repertoire_p1

repertoire_p2<-pt_summary%>%
  ggplot()+
  aes(x = `Pre-Lymphoma`, y = sf_count)+
  geom_boxplot(outlier.shape = NULL)+
  geom_beeswarm(
    aes(shape = interaction(
      Cryoglobulin,
      `B cell depletion`,
      sep = "\n+\n",
      drop = T
    )),
    alpha = 1,
    cex = 3,
    size = 3,corral.width = 1
  )+
  scale_shape_manual(values = c(20,17,18,15),name = "")+
  guides(shape = guide_none())+
  geom_pwc(bracket.nudge.y = 0.1,p.adjust.method = "bonferroni",label = "p.adj.signif", hide.ns = T)+
  scale_y_continuous(expand = expansion(add = c(1,1)))+
  scale_fill_met_d("Egypt")+
  facet_wrap(vars("VH Subfamily Count"))+
  ylab("No. Subfamilies")+
  egg::theme_article(base_size = 11)+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
  )
repertoire_p2

p_1_3_data<-pt_summary %>% 
  mutate(
    Other = ifelse(str_detect(subfamilies,"1-69"),sf_count-1,sf_count),
    `IGHV1-69` = ifelse(str_detect(subfamilies,"1-69"),1,0)) %>% 
  pivot_longer(c("Other",`IGHV1-69`),names_to = "n_sf",values_to = "value") %>% 
  group_by(`Pre-Lymphoma`,n_sf) %>% 
  summarise(n_ = sum(value)) 
p_1_3_stats = p_1_3_data%>% 
  pivot_wider(names_from = "n_sf",values_from = "n_",values_fn = as.numeric) %>% 
  column_to_rownames("Pre-Lymphoma") %>% 
  pairwise_fisher_test(p.adjust.method = "bonferroni") %>% 
  filter(p.adj < 0.05)

p_1_3<-p_1_3_data %>% 
  mutate(n_sf = factor(n_sf) %>% fct_rev()) %>% 
  arrange() %>% 
  ggplot(aes(x = `Pre-Lymphoma`,y = n_,fill = n_sf,label = n_sf))+
  geom_col(position = position_fill(),alpha = 0.8,color = "black")+
  geom_text(position = position_fill(vjust = 0.5),size = 3)+
  scale_y_continuous(labels = function(x) scales::percent(x),expand = expansion(mult = c(0,0.2)),breaks = c(0,0.25,0.50,0.75,1))+
  ylab("Proportion")+
  facet_wrap(vars("Proportion of subfamilities"))+
  scale_fill_manual(values = met.brewer(name = "VanGogh3",2))+
  egg::theme_article(base_size = 11)+
  geom_bracket(
    aes(xmin = group1, xmax = group2, label = signif(p.adj, 2)),
    data = p_1_3_stats, y.position = 1.05,inherit.aes = F,
    step.increase = 0.003,tip.length = c(0.001,0.001)
  )+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
  )
p_1_3

# Use portable default sans-serif fonts.
plot_1<-plot_grid(
  plot_grid(
    p_rf_level,
    repertoire_p1,
    labels = c("A","B"),
    rel_widths = c(5,7),
    nrow = 1),
  plot_grid(
    repertoire_p2,
    p_1_3,
    ggplot() + theme_void(),
    nrow = 1,
    rel_widths = c(5,5,2),
    labels = c("C","D","")),
  ncol = 1,
  rel_heights = c(0.7,0.5))

plot_1

ggsave("outputs/Fig1.png",plot_1,dpi = 300,width = 9.5,height = 6,bg = "white")
```

## 7. Figure 2 — mutation summaries

Summarise mutations using cohort, participant and peptide ID together.

```{r analysis-7}
shm_number <- pt_summary %>%
  ungroup() %>%
  select(`Pre-Lymphoma`, pt, sf_count, `B cell depletion`, rf_level) %>%
  left_join(
    filtered_data %>%
      count(pt, name = "shm_count"),
    by = "pt"
  ) %>%
  mutate(
    shm_count = coalesce(shm_count, 0L),
    number_of_shm_per_sf = shm_count / sf_count
  )

shm_num_total_p<-shm_number %>%
  ggplot()+
  aes(x = `Pre-Lymphoma`, y = shm_count)+
  geom_boxplot(outlier.shape = NULL,outlier.colour = "transparent")+
  geom_beeswarm(cex=2,aes(fill = `B cell depletion`))+
  scale_fill_grey()+
  egg::theme_article(base_size = 11)+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
        )+
  ylab("AA changes, N")+
  geom_pwc(group.by = "x.var",bracket.nudge.y = 0.1,hide.ns = T)+
  scale_y_continuous(expand = expansion(mult = c(0.1,0.1)))+
  facet_grid(cols = vars("SHMs Total"))
shm_num_total_p

shm_num_persf_p<-shm_number %>% 
  ggplot()+
  aes(x = `Pre-Lymphoma`, y = number_of_shm_per_sf)+
  geom_boxplot(outlier.shape = NULL,outlier.colour = "transparent")+
  geom_beeswarm(cex=2,aes(shape=`B cell depletion`),size =3)+
  egg::theme_article(base_size = 11)+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
  )+
  ylab("AA changes, N")+
  geom_pwc(group.by = "x.var",bracket.nudge.y = 0.1,hide.ns = T)+
  scale_y_continuous(expand = expansion(mult = c(0.1,0.1)))+
  facet_grid(cols = vars("SHMs per Subfamily"))
shm_num_persf_p

shm_number_1_69_only <- complete_data %>%
  ungroup() %>%
  filter(subfamily == "1-69") %>%
  distinct(`Pre-Lymphoma`, pt, sf_count, `B cell depletion`, rf_level) %>%
  left_join(
    filtered_data %>%
      filter(subfamily == "1-69") %>%
      count(pt, name = "shm_count"),
    by = "pt"
  ) %>%
  mutate(shm_count = coalesce(shm_count, 0L))

p_shm_number_1_69_only<-shm_number_1_69_only %>% 
  ggplot()+
  aes(x = `Pre-Lymphoma`, y = shm_count)+
  geom_boxplot(outlier.shape = NULL,outlier.colour = "transparent")+
  geom_beeswarm(cex = 2,aes(shape = `B cell depletion`),size = 3)+
  egg::theme_article(base_size = 11)+
  theme(legend.position = "right",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
  )+
  ylab("AA changes, N")+
  geom_pwc(group.by = "x.var",bracket.nudge.y = 0.1,hide.ns = T)+
  scale_y_continuous(expand = expansion(mult = c(0.1,0.1)))+
  facet_grid(cols = vars("IGHV1-69"))
p_shm_number_1_69_only

d_mutations_per_peptide <- results_shm %>%
  filter(coverage_count > 1) %>%
  group_by(Lymphoma, pt, peptide_unique_id) %>%
  mutate(
    mutations_per_peptide = sum(aa != ".") / n()
  ) %>%
  distinct(Lymphoma, pt, peptide_unique_id, .keep_all = TRUE) %>%
  ungroup()

p_mutations_per_peptide<-d_mutations_per_peptide %>% 
  bind_rows(d_mutations_per_peptide %>% mutate(subfamily = "all")) %>% 
  group_by(subfamily) %>% 
  filter(n_distinct(`Pre-Lymphoma`)>1) %>% 
  ggplot(aes(x = subfamily, y = mutations_per_peptide, color =  `Pre-Lymphoma`) )+
  geom_beeswarm(dodge.width = 1,corral = 'random',corral.width = 0.2,size=1,alpha=0.5)+
  geom_boxplot(position = position_dodge(width = 1),width=0.4,alpha=0.7)+
  geom_pwc(group.by = 'x.var',dodge = 1,hide.ns = T,label = "p.adj.signif",step.increase = 0.2)+
  facet_wrap(vars(subfamily),scales="free_x",nrow = 1)+
  egg::theme_article(base_size = 11)+ 
  scale_color_viridis_d(end = 0.8)+
  scale_y_continuous(expand = expansion(mult = c(0.1,0.2)))+
  ylab(str_wrap("Proportion of peptide mutated",20))+
  theme(legend.position = "right",
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 10,face = "bold"),
  )
p_mutations_per_peptide 

d_mutations_per_peptide %>% 
  ggplot()+
  aes(x = pt, y = mutations_per_peptide, color = `Pre-Lymphoma`)+
  geom_boxplot()+
  facet_wrap(vars(subfamily))
```

## 8. Figure 3 — antibody alignments and predicted affinity

Align heavy/light sequences with MUSCLE and read the five precomputed mCSM-ab2 scans.

```{r analysis-8, fig.width=20, fig.height=7}
seq_names = c(
  "07-12",
  "07-06",
  "003-27",
  "003-40",
  "YES8c",
  "Germline"
)

heavy_chain <- Biostrings::readAAStringSet("data/sequences/heavy_chain.fasta")
names(heavy_chain)<-seq_names

heavy_chain_aligned <- msa::msa(heavy_chain,method = "Muscle") %>% Biostrings::AAStringSet()

light_chain <- Biostrings::readAAStringSet("data/sequences/light_chain.fasta")

names(light_chain)<-seq_names

light_chain_aligned <- msa::msa(light_chain,method = "Muscle") %>% Biostrings::AAStringSet()

 

msa_data = list()
msa_data[["heavy_chain"]][["seq"]] = heavy_chain
msa_data[["light_chain"]][["seq"]] = light_chain
msa_data[["heavy_chain"]][["aligned"]] = heavy_chain_aligned
msa_data[["light_chain"]][["aligned"]] = light_chain_aligned

for (i in 1:length(msa_data)) {
  sequences = msa_data[[i]][["seq"]]
  chain_not_germline = sequences[names(sequences) != "Germline"]
  germline = sequences[names(sequences) == "Germline"]
  stopifnot(length(germline) == 1)
  
  germline_seq = germline[[1]] %>% as.character() %>% str_split_1("")
  
  msa_data[[i]][["mutations"]] = chain_not_germline %>% 
    as.list() %>% 
    map(
      ~{as.character(.x) %>% substr(start = 1,stop = length(germline_seq)) %>% str_split_1("") %>% {which(germline_seq != .)}}
    )
}

ddg_files <- sort(list.files("data/mcsm_ab2", pattern = "_affinity_scan[.]csv$", full.names = TRUE))
stopifnot(length(ddg_files) == 5)

for (file in 1:length(ddg_files)) {
  
  filename = ddg_files[file]
  clean_name <- sub("_affinity_scan[.]csv$", "", basename(filename))
  
  data = read_csv(filename)%>% 
    select(DDG:outcome)%>% 
    mutate(DDG = DDG)
  
  msa_data[["heavy_chain"]][["ddg"]][[clean_name]] = data %>% filter(chain == "H") 
  msa_data[["light_chain"]][["ddg"]][[clean_name]] = data %>% filter(chain == "L") 
  
  chains = names(msa_data)
  
  for (c in 1:length(chains)) {
    
    chain_cur = chains[c]
    aligned_chain_seq = msa_data[[chain_cur]][["aligned"]][[clean_name]] %>% as.character()
    alignment_breaks = {gregexpr(pattern ='-',aligned_chain_seq)[[1]]} %>% as.numeric() %>% base::sort(decreasing = F)
    current_ddg_table = {msa_data[[chain_cur]][["ddg"]][[clean_name]]} 
    current_ddg_table = current_ddg_table%>% mutate(position_aligned = position)
    
    for(b in 1:length(alignment_breaks)) {
      a_break = alignment_breaks[b]
      
      current_ddg_table = current_ddg_table %>% 
        mutate(position_aligned = case_when(
          a_break == -1 ~ position_aligned,
          position_aligned >= a_break ~ position_aligned +1,
          TRUE ~ position_aligned
        ))
    }
    msa_data[[chain_cur]][["ddg_aligned"]][[clean_name]] = current_ddg_table
    
    msa_data[[chain_cur]][["ddg_summary"]][[clean_name]] = current_ddg_table %>% 
      group_by(position,position_aligned) %>% 
      summarise(DDG_mean = mean(DDG),DDG_median = median(DDG))
    
    msa_data[[chain_cur]][["contacts"]][[clean_name]] = current_ddg_table$position_aligned %>% 
      unique() %>% 
      sort()
    
    mutations = msa_data[[chain_cur]][["mutations"]][[clean_name]]
    chain_seqs = msa_data[[chain_cur]][["seq"]]
    chain_seqs_s = chain_seqs[names(chain_seqs) == "Germline"] %>% 
      as.character() %>% 
      str_split_1("")
    
    msa_data[[chain_cur]][["mut_ddg"]][[clean_name]] = tibble(
      chain_c = chain_cur,
      position = mutations,
      mutant = chain_seqs_s[mutations]
    ) %>% 
      left_join(msa_data[[chain_cur]][["ddg_aligned"]][[clean_name]]) %>% 
      filter(!is.na(DDG))
    
  }
}

msa_graph_data = list(
  hc = list(
    sequence = msa_data[["heavy_chain"]][["aligned"]],
    cutpoint = list(c(25,33),c(50,58),c(96,113)),
    mut_ddg  = msa_data[["heavy_chain"]][["mut_ddg"]], 
    contacts = msa_data[["heavy_chain"]][["contacts"]],
    sum_ddg = msa_data[["heavy_chain"]][["ddg_summary"]]
  ),
  lc = list(
    sequence = msa_data[["light_chain"]][["aligned"]],
    cutpoint = list(c(26,33),c(50,53),c(89,98)),
    mut_ddg  = msa_data[["light_chain"]][["mut_ddg"]], 
    contacts = msa_data[["light_chain"]][["contacts"]],
    sum_ddg = msa_data[["light_chain"]][["ddg_summary"]]
  )
)

make_msa_graph<-function(msa_cdr_list){
  out_graph_list <- list()
  for (chain in 1:length(msa_cdr_list)) {
    sequence = msa_cdr_list[[chain]][["sequence"]]
    CDR_cuts = msa_cdr_list[[chain]][["cutpoint"]]
    contact_resi = msa_cdr_list[[chain]][["contacts"]]
    sum_ddg = msa_cdr_list[[chain]][["sum_ddg"]]
    spec_ddg = msa_cdr_list[[chain]][["mut_ddg"]]
    
    p_msa<-ggmsa(
      sequence,
      consensus_views = T,
      seq_name = T,
      by_conservation = F,
      ref = "Germline",
      color = "Zappo_AA",
      char_width = 0.7, 
      border = "transparent",
      use_dot = T
    )+
      egg::theme_article(base_size = 11)+
      scale_x_continuous(expand = expansion(add = c(0.5,2)),breaks = scales::breaks_width(5))+
      scale_y_discrete(expand = expansion(add = c(0,3.5)))
    
    for (i in 1:length(msa_cdr_list[[chain]][["sequence"]])) {
      data_ = tibble(
        x_ = seq.int(
          1,
          length(msa_cdr_list[[chain]][["sequence"]][[i]]))
      ) %>% 
        mutate(y_ = names(msa_cdr_list[[chain]][["sequence"]][i]))
      
      p_msa<-p_msa +
        geom_tile(
          data = data_,
          aes_(x = data_$x_, y = data_$y_),
          inherit.aes = F,
          linewidth = 0.6, 
          color = "white", fill = "transparent",width = 1, height = 1)
    }
    
    for (i in 1:length(CDR_cuts)) {
      
      p_msa<-p_msa +
        geom_vline(xintercept = min(CDR_cuts[[i]])+0.5,linewidth=0.5,linetype = 2)+
        geom_vline(xintercept = max(CDR_cuts[[i]])+0.5,linewidth=0.5,linetype = 2)+
        geom_rect(fill = "grey20",
                  aes_(
                    xmin = min(CDR_cuts[[i]])+0.5,
                    xmax = max(CDR_cuts[[i]])+0.5,
                    ymin = -1,
                    ymax = 6), alpha = 0.1
        )+
        geom_text(aes_(
          x = mean(CDR_cuts[[i]])+0.5,
          y = 8,
          label = paste0("CDR",i)
        ))
    }
    
    for (i in 1:length(contact_resi)) {
      name_ = contact_resi[i] %>% names()
      msa_temp_data = tibble(
        position_aligned = contact_resi[[i]]
      ) %>% 
        mutate(
          name = name_
        ) %>% 
        left_join(spec_ddg[[name_]])
      
      p_msa<-p_msa +
        geom_tile(
          aes_(
            x = msa_temp_data$position_aligned,
            y = msa_temp_data$name,
            color = msa_temp_data$DDG
          ),
          fill = "transparent",
          width = 0.8, height = 0.1, position = position_nudge(y=-0.5),
          linewidth = 1,
          linetype = 1,
          inherit.aes = F)+
        scale_color_gradient2(
          low = "red",
          mid = "grey70",
          midpoint = 0,
          high = "blue",
          na.value = "black",
          name = "ddG",
          breaks = c(-2,0,2),
          limits = c(-2,2),
          oob = scales::squish,
          minor_breaks = NULL,
          labels = c("Reduce","0", "Increase"),
        )+
        theme(legend.text = element_text(vjust = 0.5),legend.title = element_text(vjust = 0.8))
    }
    sum_ddg_comb = tibble()
    for (d in 1:length(sum_ddg)) {
      sum_ddg_comb = sum_ddg_comb %>% bind_rows(
        sum_ddg[[d]] %>% mutate(name = names(sum_ddg[d]))
      )
    }
    
    ddg_comb = sum_ddg_comb %>% 
      group_by(position_aligned) %>% 
      filter(!is.na(DDG_mean)) %>% 
      summarise(ddg_mean_all = mean(DDG_mean))
    
    p_msa<-p_msa +
      geom_tile(
        aes_(
          x = ddg_comb$position_aligned,
          y = 0,
          fill = ddg_comb$ddg_mean_all
        ),
        color = "black",
        width = 0.9, 
        height = 0.9,
        linewidth = 0.1,
        linetype = 1,
        inherit.aes = F,
        show.legend = F,position = position_nudge(y = -0.5))+
      scale_fill_gradient2(
        low = "red",
        mid = "grey70",
        midpoint = 0,
        high = "blue",
        na.value = "black",
        name = "ddG",
        breaks = c(-2,0,2),
        oob = scales::squish,
        limits = c(-2,2),
        minor_breaks = NULL,
        labels = c("Reduce","0", "Increase")
      )
    
    
    out_graph_list[[chain]]<-p_msa
    
  }
  
  ggpubr::ggarrange(plotlist = out_graph_list,ncol = 1,labels = LETTERS,common.legend = T)
}

make_msa_graph(msa_graph_data)->msa_graph
msa_graph
ggsave2("outputs/Fig3.png",msa_graph,bg = "white",dpi = 600,width = 20,height = 7)
```

## 9. Figure 2 — hotspot panel and assembly

Panel D shows substitutions observed in more than two participants across the lymphoma groups; asterisks mark YES8c contacts.

```{r analysis-9}
d_mutation_hot_spots <- filtered_data %>%
  filter(subfamily == "1-69") %>%
  group_by(aa_position, aa, Lymphoma, aa_ref) %>%
  summarise(count = n_distinct(pt)) %>%
  group_by(aa_position,aa) %>% 
  filter(sum(count)> 2) %>%
  mutate(position = case_when(
    Lymphoma == "Lymphoma" ~ count+1,
    TRUE ~ -count-1
  ))%>% 
  left_join(aa_zappo, join_by(aa)) %>%
  arrange(aa_position) %>%
  left_join(aa_zappo %>% 
              select(group, aa, color) %>% 
              dplyr::rename(c(
                ref_group = "group", 
                ref_color = "color")),
            join_by(aa_ref == aa)) %>% 
  mutate(label=as_factor(paste0(aa_ref,fct_inseq(fct(as.character(aa_position))))))

reference_labels=d_mutation_hot_spots %>% 
  ungroup() %>% 
  select(aa_ref,ref_group,label) %>% 
  mutate(position = 1,Lymphoma = 'Reference') %>% 
  distinct(label,.keep_all = T)

contacts = msa_data$heavy_chain$contacts$YES8c

plot2_mutation_hot_spots=d_mutation_hot_spots %>% 
  ggplot(aes(
    y = position,
    x = label,
    label = paste0(aa, ": ", abs(count)),
    fill = group,
    group = aa
  )) +
  geom_col(
    size = 10,
    position = position_dodge(),
    color = "black",
    linewidth = 0.5
  ) +
  geom_text(aes(hjust = ifelse(Lymphoma=="Lymphoma",-0.5,1.5)),color = "black", position = position_dodge(width = 0.8),size=3,angle = 90) +
  geom_text(data=data.frame(),aes(x = 3, y = 10, label = "Lymphoma"),inherit.aes = F,size = 4)+
  geom_text(data=data.frame(),aes(x = 3, y = -10, label = "Non-Lymphoma"),inherit.aes = F,size = 4)+
  geom_box(inherit.aes = F,show.legend = F,
           data = reference_labels,
           linewidth = 1,
           color = "black",
           aes(
             ymin = position,
             ymax = -position,
             x = label,
             fill = ref_group,
             width = 0.9,
           )) +
  geom_text(data = reference_labels,
            aes(label = ifelse(str_extract(label,"\\d+") %>% as.numeric() %in% msa_data$heavy_chain$contacts$YES8c, paste0(label,"*"), paste0(label)),
                y = 0,x=label),size = 3,inherit.aes = F)+
  scale_y_continuous(labels = abs,expand = expansion(add = 3)) +
  scale_zappo +
  theme_pubr(legend = "bottom",base_size = 10) +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 8),
    panel.border = element_rect(color = "black", size = 0.5,fill = "transparent")
    )+
  guides(fill = guide_legend(nrow = 1),
         y = guide_axis_truncated(
           trunc_lower = c(-Inf,1),
           trunc_upper = c(1,Inf)
         ))+
  ylab("") + xlab("AA number(sequential)")
plot2_mutation_hot_spots

plot_grid(
  plot_grid(
    plot_grid(
      shm_num_persf_p,
      p_shm_number_1_69_only,
      labels = c("A","B"),
      align = "h",
      axis = "lr",
      nrow = 1,
      rel_widths = c(0.8, 1)
    ),
    plot_grid(
      p_mutations_per_peptide,
      labels = "C"
    ),
    ncol = 1
  ),
  plot2_mutation_hot_spots,
  labels = c("", "D"),
  align = "",
  axis = "tb",
  ncol = 1,
  rel_heights = c(1, 0.8)
  
) -> plot_2

plot_2

ggsave2("outputs/Fig2.png",plot_2,bg = "white"
      ,dpi = 600,width = 10,height = 8)
```

## 10. Figure 4A — interface-associated mutations

Compare mutation counts using the union of heavy-chain contacts from five models. Figure 4B is not implemented in the supplied script.

```{r analysis-10}
extended_contacts = tibble()
all_contacts = msa_graph_data$hc$contacts
for (d in 1:length(all_contacts)) {
  extended_contacts = extended_contacts %>% bind_rows(tibble(
    contact = all_contacts[[d]]
  ) %>% mutate(name = names(all_contacts[d]))
  )
}

interface_counts <- filtered_data %>%
  filter(subfamily == "1-69") %>%
  mutate(
    interfacial = if_else(
      aa_position %in% unique(extended_contacts$contact),
      "Antigen binding",
      "Non-interfacing"
    )
  ) %>%
  count(pt, interfacial, name = "count")

interface_data <- tidyr::crossing(
  complete_data %>%
    ungroup() %>%
    filter(subfamily == "1-69") %>%
    distinct(`Pre-Lymphoma`, pt),
  interfacial = c("Antigen binding", "Non-interfacing")
) %>%
  left_join(interface_counts, by = c("pt", "interfacial")) %>%
  mutate(count = coalesce(count, 0L))

fig_4 <- interface_data %>%
  ggplot() +
  aes(x = interfacial,y = count, color = `Pre-Lymphoma`)+
  geom_boxplot(position = position_dodge(width = 0.9))+
  geom_beeswarm(dodge.width = 0.9,size = 2,cex = 2)+
  scale_color_viridis_d(end = 0.8)+
  geom_pwc(p.adjust.method = "none", hide.ns = T)+
  egg::theme_article(base_size = 11) +
  ylab("Mutation Count")+
  theme(
    strip.text = element_text(face = "bold"),
    legend.text = element_text(size = 8),
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )+
  facet_wrap(vars("VH1-69 RF mutated residues"))
fig_4
# Figure 4B structural panels are not defined by the supplied script.
ggsave("outputs/Fig4A.png", bg = "white", fig_4, width = 9, height = 4, dpi = 300)
```

## 11. Supplementary analyses

Compare mutation properties and the 23 supplied HCDR3 sequences.

```{r analysis-11, fig.width=12, fig.height=9}
##### SUPP 1: Charge and Hydro####

hydro_p<-filtered_data %>%
  ggplot()+
  aes(x = `Pre-Lymphoma`,fill = Hydrophobic_change,label = Hydrophobic_change)+
  geom_bar(position = position_fill())+
  scale_y_continuous(labels = function(x) percent(x))+
  theme_pubr()+
  theme(legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 10,face = "bold"),
  )+
  ylab("Proportion")+
  facet_grid(cols = vars("Hydrophobicity change"))+
  scale_fill_met_d("Egypt",direction = -1)
hydro_p

charge_p<-filtered_data %>% 
  filter(subfamily == "1-69") %>%
  group_by(`Pre-Lymphoma`,Charge_change) %>% 
  mutate(label = n()) %>% 
  ggplot()+
  aes(x = `Pre-Lymphoma`,fill = Charge_change,label = label)+
  geom_bar(position = position_fill())+
  scale_y_continuous(labels = function(x) percent(x))+
  theme_pubr()+
  theme(legend.position = "bottom", legend.title = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 8),
        strip.text = element_text(size = 10,face = "bold"),
  )+
  coord_flip()+
  facet_grid(cols = vars("Charge Change"))+
  scale_fill_met_d("Hiroshige")
charge_p

hydro_p <- hydro_p + labs(fill = NULL)
supp1 <- ggarrange(hydro_p,charge_p,nrow = 1, labels = LETTERS)
supp1
ggsave("outputs/Supplementary1.png", bg = "white", supp1, width = 10, height = 5, dpi = 300)

filtered_data %>% 
  group_by(Lymphoma) %>% 
  mutate(change_type = case_when(
    Hydrophobic_change == "Hydrophilic" ~ "Yes",
    TRUE ~ "No"
  ))->hydrophob_shm_stat

ftable(hydrophob_shm_stat$Lymphoma,hydrophob_shm_stat$change_type) %>%
  fisher_test() %>%
  as_tibble() %>% 
  {write_csv(., "outputs/hydrophobicity_fisher_test.csv"); knitr::kable(.)}

####Supplementary 2: HCDR3s ####

collated_HCDRs <- Biostrings::readAAStringSet("data/sequences/hcdr3_comparison.fasta")

collated_HCDRs[names(collated_HCDRs) %>% str_detect("\\*")]
collated_HCDRs[names(collated_HCDRs) %>% str_detect("\\*",negate = T)]
msa::msaClustalW(collated_HCDRs) %>% Biostrings::AAMultipleAlignment()->msa_misc_hcdr
supp2 <- ggmsa::ggmsa(msa_misc_hcdr,
             color = "Zappo_AA",
             consensus_views = T,
             by_conservation = F,
             seq_name = T,
             ignore_gaps = F,
)+
  geom_seqlogo()+
  theme_pubr()+
  scale_x_continuous(breaks = c(0,4,8,12,16,20),expand = expansion(add = c(0,0)))+
  theme(axis.line.y = element_blank())

supp2
ggsave("outputs/Supplementary2.png", bg = "white", supp2, width = 12, height = 9, dpi = 300)
```

## 12. Export data, statistics and reproducibility diagnostics

CSV exports retain participant codes as text. When opening them in a spreadsheet,
import identifier columns as text to avoid losing leading zeros. Statistical
annotations are computed when the plots are built; their layer tables are exported
separately, including non-significant tests hidden in the displayed figures.
Bracket layer tables repeat each test for its three drawn segments; deduplicate
by panel, group pair and p-value when using these as statistical summaries.

```{r export-data}
tables <- list(participant_summary = pt_summary, parsed_residues = results_shm,
  mutation_changes = shm_changes, complete_data = complete_data,
  filtered_data = filtered_data, figure1_proportions = p_1_3_data,
  figure1_significant_fisher_tests = p_1_3_stats,
  figure2_shm_number = shm_number, figure2_ighv169 = shm_number_1_69_only,
  figure2_peptide_proportions = d_mutations_per_peptide,
  figure2_hotspots = d_mutation_hot_spots,
  interface_contacts = extended_contacts)
for (nm in names(tables)) write_csv(tables[[nm]], file.path("outputs", paste0(nm, ".csv")))
all_fisher_tests <- p_1_3_data %>%
  pivot_wider(names_from = n_sf, values_from = n_, values_fn = as.numeric) %>%
  column_to_rownames("Pre-Lymphoma") %>%
  pairwise_fisher_test(p.adjust.method = "bonferroni")
write_csv(all_fisher_tests, "outputs/figure1_all_fisher_tests.csv")
for (ch in names(msa_data)) {
  Biostrings::writeXStringSet(msa_data[[ch]]$aligned, file.path("outputs", paste0(ch, "_aligned.fasta")))
  for (nm in names(msa_data[[ch]]$ddg_aligned))
    write_csv(msa_data[[ch]]$ddg_aligned[[nm]], file.path("outputs", paste0(ch, "_", nm, "_ddg.csv")))
}
plots <- list(rf_level = p_rf_level, subfamily_count = repertoire_p2,
  shm_per_subfamily = shm_num_persf_p, shm_ighv169 = p_shm_number_1_69_only,
  peptide_proportion = p_mutations_per_peptide, interface = fig_4)
for (nm in names(plots)) {
  # Export non-significant tests too, without changing the displayed figures.
  export_plot <- plots[[nm]]
  export_plot$layers <- lapply(export_plot$layers, function(layer) {
    params <- layer$stat_params
    if (!is.null(params$hide.ns)) params$hide.ns <- FALSE
    ggplot2::ggproto(NULL, layer, stat_params = params)
  })
  layers <- ggplot_build(export_plot)$data
  for (j in seq_along(layers)) {
    if (any(grepl("^p($|[.])", names(layers[[j]]))))
      write_csv(layers[[j]], file.path("outputs", paste0("statistics_", nm, "_layer", j, ".csv")))
  }
}
# A short ID may recur across cohorts; only the full key identifies a peptide.
id_audit <- results_shm %>% distinct(peptide_unique_id, pt, Lymphoma) %>%
  count(peptide_unique_id, name = "participant_cohort_combinations") %>%
  filter(participant_cohort_combinations > 1)
write_csv(id_audit, "outputs/reused_local_peptide_ids.csv")
composite_audit <- d_mutations_per_peptide %>%
  count(Lymphoma, pt, peptide_unique_id, name = "summary_rows") %>%
  filter(summary_rows > 1)
write_csv(composite_audit, "outputs/duplicate_composite_peptide_keys.csv")
threshold_audit <- filtered_data %>%
  filter(is.na(specific_aa_mut_coverage) | is.na(specific_mut_proportion) |
           specific_aa_mut_coverage <= 1 | specific_mut_proportion <= 0.1)
write_csv(threshold_audit, "outputs/nonqualifying_retained_rows.csv")
stopifnot(nrow(composite_audit) == 0L, nrow(threshold_audit) == 0L,
          nrow(shm_number) == nrow(pt_summary))
audit <- tibble(metric = c("Participants", "Parsed residue observations",
  "Qualifying mutation rows", "Nonqualifying retained rows",
  "Reused local peptide IDs (expected; not composite-key collisions)",
  "Peptide summaries using composite keys", "Duplicate composite peptide keys"),
  value = c(nrow(pt_summary), nrow(results_shm), nrow(filtered_data),
            nrow(threshold_audit), nrow(id_audit), nrow(d_mutations_per_peptide),
            nrow(composite_audit)))
write_csv(audit, "outputs/run_summary.csv")
knitr::kable(audit)
writeLines(capture.output(sessionInfo()), "outputs/session_info.txt")
saveRDS(tables, "outputs/analysis_tables.rds")
```

## Notes

RF units are IU/L. Peptide-level tests do not model within-participant clustering.
Supplemental counts distinguish unfiltered records from qualifying mutations.
Setup and data terms are in `README.md`.

```{r session-info}
sessionInfo()
```
